Fast algebraic filtering of surfaces from 3D medical images with Julia

### 2019-05-09 src/filter.tex
update filter.tex file

For the sake of simplicity, we assume a common size on the three image axes, and the corresponding image portion, called \emph{block} $B$, as a function of its element of the  lowest single block  coordinates $i,j,k\in\N$ and block dimension $n$:
\[
B(i,j,k,n) := Image([in,jn,kn],[i(n+1),j(n+1),k(n+1)]) 
\]

@@

Figure~\ref{fig:blocks} shows the block decomposition in a 2D case, with positive integer $u,v,w$ lateral sizes of image. Note that block sides do not necessarily correspond to image edges.

@@

>>>>

For the sake of simplicity, we assume a common size on the three image axes, and the corresponding image portion $B$, called \emph{block}, as a function of its element of the  lowest  block  coordinates $i,j,k\in\N$ and of block dimension $n$:
\[
B(i,j,k,n) := Image([in,jn,kn],[i(n+1),j(n+1),k(n+1)]) 
\]

@@

Figure~\ref{fig:blocks} shows the block decomposition in a 2D case, with positive integer $u,v$ lateral sizes of image. Note that block sides do not necessarily correspond to image edges.

@@

Here we refer directly to the 3D case.
Let us call \emph{segment} the bulk content $S$ of interest within the input 3D image of size $(u,v,w)$. We aim to compute the segment boundary $\partial_3 S$. 
First we set the size $n$ of the block, in order to decompose the input $Image(u,v,w)$ into a fair number 
\[
M = \ceil{u/n} \times \ceil{v/n} \times \ceil{w/n}
\] of blocks. 
Then we consider each image portion $c_{i,j,k} = S\cap B(i,j,k,n)$ and compute its (binary) coordinate representation  $[c]_{i,j,k}\in C_3(n,n,n)$. This one is a sparse binary vector of length $n^3$. Then assemble the $M$ representations $c$ of segment portions into a sparse binary matrix $\T{S}$, of dimension $n^d \times M$. Finally compute a matrix $\T{B}$ of boundary portions of $S$, represented by columns as chain coordinate vectors in $C_2$:
\[
\T{B} = [\partial_3(n,n,n)]\, \T{S}.
\]

Of course, the $\T{B}$ sparse matrix has the same column number $M$ of $\T{S}$, because each column contains the boundary representation of the corresponding $S\cap B_{i,j,k}$, and the number of rows of the operator, equal to the dimension of the linear space $C_2$.

\paragraph{Embedding}
A final computational step is needed, in order to embed the 2-chains in Euclidean space $\E^3$ and to assemble the whole resulting surface. In particular, we need to compute the \emph{embedding function} $\mu : C_0 \to \E^3$, where $C_0$ is the space of 0-chains, one-to-one corresponding to the vertices of the extracted surface. The simplest solution is to associate  four 0-cells to each 2-cell of the extracted surface, i.e.~to each non-zero entry in every column of \T{B}.  The $\mu$ function  can be computed by identifying, via  element position in the column, a triple of integer values $0\leq x\leq u, 0\leq y\leq v, 0\leq z\leq w$ for each vertex of the 2-cell.  The mapping can be implemented using a dictionary, thet will store the inverse coordinate transformation used at the beginning, i.e.~the one from linear to Cartesian coords, in order of not duplicating the output vertices.

\paragraph{Assembling the surface portions}
